---
title: "Biological sequences in tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gtseq)
```


## Example

gtseq provides several functions to help create `gt` tables of
biological sequences. Although the philosophy is to provide small
modular functions to allow for high customizability, `gtseq::gtseq` is a
wrapper function provided to quickly and efficiently explore sequences.

Use `gtseq` on a `data.frame` with a column containing amino acid
sequences:

```{r}
library(gtseq)

msa |> gtseq(seq_column = seq)
```

For further customization use the lower-level functions:

## `split_sequences`

`split_sequences` for structuring the data for `gt`, the output of this
function retains the 1 sequence per row structure, but splits each
element of the sequence data, for maximum flexibility for customization
in `gt`

```{r}
library(gt)
msa |> split_sequences(seq_column = seq)
```

## `apply_color_to_aa`

`apply_color_to_aa` is for coloring amino acids by one of the included
palettes

A function is returned, which takes an input and colors according to a
given palette

```{r}
apply_color_to_aa("Chemistry")(c("A", "K", "L", "-"))
```

This function can be directly passed to `gt::data_color`

```{r}
msa |>
  split_sequences(seq_column = seq) |>
  gt() |>
  data_color(columns = contains('pos_'), fn = apply_color_to_aa('Chemistry'))
```

## `get_consensus`

`get_consensus` for adding consensus sequence and frequency data. Pass
to either `gt::grand_summary_rows` or `gt::summary_rows`.

```{r}
msa |>
  split_sequences(seq_column = seq) |>
  gt() |>
  data_color(columns = contains('pos_'), fn = apply_color_to_aa('Chemistry')) |>
  grand_summary_rows(
    columns = contains('pos_'),
    fns = list(
      consensus = ~ get_consensus(.)
    )
  )
```

## `generate_breaks`

Takes a vector and returns every nth element, based on
`scales::breaks_width`

```{r}
library(dplyr)
cnames <- msa |>
  split_sequences(seq_column = seq) |>
  select(contains('pos_')) |>
  colnames()
breaks <- generate_breaks(cnames)
breaks
```

Which allows only applying a label to every nth position of the
sequences with `gt::cols_label_with`

```{r}
msa |>
  split_sequences(seq_column = seq) |>
  gt() |>
  data_color(columns = contains('pos_'), fn = apply_color_to_aa('Chemistry')) |>
  cols_label_with(
    columns = contains('pos_'),
    fn = ~ ifelse(. %in% breaks, gsub('pos_', '', .), '')
  )
```
